---
title: "Automation of lab data"
author: "Ram Barankin"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##  The problem
A master's student have told me that her adviser is spending $10k (!) a year for hiring students to organize excel data. They have been using a lab tool that analyzes chemical samples (I don't have any idea of what), and the relevant part (for them) of the output of the analysis is located in the 21st sheet of an Excel file that is saved inside a separate folder for each sample (!).
They are manually doing two things:

1. Going into each folder of each sample, going to the last sheet and copying and pasting the relevant columns to a joint sheet in a different Excel file. Each sample file was named MF<sample number>.D.xls (e.g., MF744.D).

2. Organizing the data according to different categories (that are the rounded value of two digits of one of the relevant variables).
The PI apparently stated several time that experts (?) told her that this process cannot be automated.

### Challenge accepted 
At first I thought I am missing something, but I soon to find out that I can easily automate the process in R by creating the following function:


```{r 1}
#Define the function:
chels_fun <- function(start, end, path){
  #install relevant packages if needed:
  list.of.packages <- c("dplyr", "gdata")
  new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
  if(length(new.packages)) install.packages(new.packages)
  #load packages:
  library(dplyr)
  library(gdata)
  #load the first sheet:
  file <- paste0(path, "MF", start, ".D/REPORT01.xls")
  jo <- tbl_df(read.xls(file, 21))
  jo <- jo%>% 
    #create the categories: 
    mutate(Time_cat=round(RetTime,2))%>%
    dplyr::select(Time_cat, RetTime, Area)
  colnames(jo) <- c("Time_cat", paste0("Area.",start), paste0("RetTime.", start))
  # Do the same for the rest of the samples:
  a=start+1
  for (i in a:end){
    file <- paste0(path, "MF",i, ".D/REPORT01.xls")
    dat <- tbl_df(read.xls(file, 21))
    dat <- dat%>% 
      mutate(Time_cat=round(RetTime,2))%>%
      dplyr::select(Time_cat, RetTime, Area)
    #join to the rest of the samples
    colnames(dat) <- c("Time_cat", paste0("Area.",i), paste0("RetTime.", i))
    jo <- full_join(jo, dat, by="Time_cat")
  }
  
  jo <- arrange(jo, Time_cat)
}

```

Then, all my friend is needed to do is define the samples number range and the path to where the samples files are located and she gets a data frame that has all she needed.

```{r 2, message=F, warning=F}
source("chels_fun.R") #load the function
start<-744 #the first sample number
end<-747 #the last sample number
path <-("Sample data/")# the location of the sample folders (not including the sample folders themselfs!)
dat <- chels_fun(start,end,path)
dat #print the results (this will print only the first 10 lines, since it is a dplyr tbl_df data frame)
write.csv(dat,"my_results.csv") #save the results 
```

* Find the relevant data at https://github.com/RamBarankin/Automating-data-mining-of-lab-data.git